Predicting Ambient Air Pollution (PM2.5) Across the Contiguous U.S.

### 2023-04-22 Draft.Rmd
updated knn

# Create standardize train data for kNN and PCA
# Omit value and id variables since these should not be scaled
train_scaled = train[,3:50] %>% dplyr::select(where(is.numeric)) %>% mutate(across(fips:aod, scale)) 
#%>% cbind(state=train$state,county=train$county,city=train$city)

@@

```{r kNN}
set.seed(123)
# Find ideal k neighbors
sqrt(nrow(train))

# Scale the training data for k-NN model in wrangling
train_scaled = train[,3:50] %>% dplyr::select(where(is.numeric)) %>% mutate(across(fips:aod, scale)) 
#%>% cbind(state=train$state,county=train$county,city=train$city)

@@

# Create workflow
  wf2 <- workflow() %>% 
    add_recipe(rec) %>% 
    add_model(model2)

# Fit the model on the training dataset using the `fit()` function
res2 <- fit(wf2, data = train_scaled)

@@

# Check performance using cross-validation

folds2 <- vfold_cv(train, v = 10)
# folds
res2 <- fit_resamples(wf2, resamples = folds2) # ERROR Assigned data `orig_rows` must be compatible with existing data.
res2 %>% 
    collect_metrics()
```

@@

```{r Xgboost}
set.seed(123)
# Preprocessing "recipe"
preprocessing_recipe <- 
  recipes::recipe(value~lon+lat, data = training(dat_split)) %>%
  # Convert categorical variables to factors
  recipes::step_string2factor(all_nominal()) %>%
  # Combine low frequency factor levels
  recipes::step_other(all_nominal(), threshold = 0.01) %>%
  # Remove no variance predictors which provide no predictive information 
  recipes::step_nzv(all_nominal()) %>%
  prep()

@@

knitr::kable(head(xgboost_grid))

@@

# Hyperparameter tuning
# Takes a long time to run
xgboost_tuned <- tune::tune_grid(object = xgboost_wf, resamples = folds3,
  grid = xgboost_grid, metrics = yardstick::metric_set(rmse, rsq, mae),
  control = tune::control_grid(verbose = TRUE))

# Find hyper parameter values which performed best at minimizing RMSE
xgboost_tuned %>% tune::show_best(metric = "rmse") %>%knitr::kable()
# Isolate the best performing hyperparameter values.
xgboost_best_params <- xgboost_tuned %>% tune::select_best("rmse")
knitr::kable(xgboost_best_params)

@@

train_prediction <- xgboost_model_final %>%
  # Fit the model on all the training data
  fit(formula = value ~ lon+lat, data = train_processed) %>%
  # Predict the sale prices for the training data
  predict(new_data = train_processed) %>%
  bind_cols(training(dat_split))

xgboost_score_train <- train_prediction %>%
  yardstick::metrics(value, .pred) %>%
  mutate(.estimate = format(round(.estimate, 2), big.mark = ","))
knitr::kable(xgboost_score_train)
```

```{r Random Forest}
set.seed(123)
# Create Random Forest Model
set.seed(123)
# Use grid of tuning parameters
model4 <- rand_forest(mtry = tune("mtry"),
                     min_n = tune("min_n")) %>% 
    set_engine("ranger") %>% 
    set_mode("regression")
# Create workflow
  wf4 <- workflow() %>% 
    add_recipe(rec) %>% 
    add_model(model4)

>>>>

# Create standardize train data for kNN and PCA
# Omit value and id variables since these should not be scaled
# train_scaled for PCA
train_scaled = train[,3:50] %>% dplyr::select(where(is.numeric)) %>% mutate(across(fips:aod, scale)) 
#%>% cbind(state=train$state,county=train$county,city=train$city)

# train_knn for kNN model
train_knn = train[,2:50] %>% dplyr::select(where(is.numeric)) %>% mutate(across(fips:aod, scale)) %>% cbind(state=train$state,county=train$county,city=train$city)

@@

```{r kNN ALL MODELS FAILED}
set.seed(123)
# Find ideal k neighbors
sqrt(nrow(train))

# Create recipe
rec2 = train_knn |> recipe(value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+log_nei_2008_pm10_sum_10000)

@@

# Create workflow
  wf2 <- workflow() %>% 
    add_recipe(rec2) %>% 
    add_model(model2)

# Fit the model on the training dataset using the `fit()` function
res2 <- parsnip::fit(wf2, data = train_knn)

@@

# Check performance using cross-validation
folds2 <- vfold_cv(train, v = 10)
# folds
res2 <- fit_resamples(wf2, resamples = folds2) # ERROR Assigned data `orig_rows` must be compatible with existing data.
res2 %>% 
    collect_metrics()
```

@@

```{r Xgboost}
set.seed(123)
# Preprocessing "recipe"
preprocessing_recipe <- 
  recipes::recipe(value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+log_nei_2008_pm10_sum_10000, data = training(dat_split)) %>%
  # Convert categorical variables to factors
  recipes::step_string2factor(all_nominal()) %>%
  # Combine low frequency factor levels
  recipes::step_other(all_nominal(), threshold = 0.01) %>%
  # Remove no variance predictors which provide no predictive information 
  recipes::step_nzv(all_nominal()) %>%
  prep()

@@

head(xgboost_grid)

@@

# Hyperparameter tuning
# Takes a long time to run, 5 min
xgboost_tuned <- tune::tune_grid(object = xgboost_wf, resamples = folds3,
  grid = xgboost_grid, metrics = yardstick::metric_set(rmse, rsq),
  control = tune::control_grid(verbose = TRUE))

# Find hyper parameter values which performed best at minimizing RMSE
xgboost_tuned %>% tune::show_best(metric = "rmse") %>%knitr::kable()
# Isolate the best performing hyperparameter values.
xgboost_best_params <- xgboost_tuned %>% tune::select_best("rmse")
xgboost_best_params

@@

train_prediction <- xgboost_model_final %>%
  # Fit the model on all the training data
  parsnip::fit(formula = value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+log_nei_2008_pm10_sum_10000, data = train_processed) %>%
  # Predict the sale prices for the training data
  predict(new_data = train_processed) %>%
  bind_cols(training(dat_split))

xgboost_score_train <- train_prediction %>%
  yardstick::metrics(value, .pred) %>%
  mutate(.estimate = format(round(.estimate, 2), big.mark = ","))
xgboost_score_train
```

```{r Random Forest}
set.seed(123)
# Create Random Forest Model
# Use grid of tuning parameters
model4 <- rand_forest(mtry = tune("mtry"),
                     min_n = tune("min_n")) %>% 
    set_engine("ranger") %>% 
    set_mode("regression")
# Create workflow
  wf4 <- workflow() %>% 
    add_recipe(rec) %>% 
    add_model(model4)
